---
title: "DrugB code"
output: pdf_document
---
### 1) Importing XML and exploring it:
## Setting up the directory and the xml file :

```{r}
path=c("Drugbank data/")
drugbank.xml <- "database.xml"
```

## Parse the xml file :
```{r}
#install.packages("XML")#Don't run if already installed, if not remove the #
library(XML)
library(plyr)
library(parallel)
xmlfile <- xmlParse(paste0(path, drugbank.xml))
xmltop <- xmlRoot(xmlfile) #gives content of root
```

## Explore it :
An Xml file has a tree structure. It starts at a root element and branches into child elements. Example : 

<root>
  <child>
    <subchild>.....</subchild>
  </child>
</root>


Let's see the structure of our drugbank.xml file:
```{r}
class(xmlfile) #"XMLInternalDocument" "XMLAbstractDocument"
class(xmltop) # "XMLInternalElementNode" "XMLInternalNode" "XMLAbstractNode"
xmlName(xmltop) # give name of node: "drugbank"
xmlSize(xmltop) # how many children in node, you ghave to find 11292
xmlName(xmltop[[1]]) # name of root's children : "drug"
```
Now we can extract entries, for example number k=325:
```{r}
k <- 325
xmltop[[k]][["drugbank-id"]] #The drug's ID is DB00335
xmltop[[k]][["name"]] # The drug's name is Atenolol
xmltop[[k]][["indication"]][[1]]#For what the drug is used
xmltop[[k]][["description"]]# Description of the drug 
xmltop[[k]][["snp-effects"]]#Target protein name, gene name, uniprot id, SNP id...
xmltop[[k]][["snp-adverse-drug-reactions"]]#No SNP adverse drug reaction
xmltop[[k]][["categories"]][[2]]#Category of the drug: Adrenergic Antagonists
xmltop[[k]][["targets"]][[1]][[7]][["gene-name"]]#Gene of the target protein
xmltop[[k]][["carriers"]][[1]][[7]][["gene-name"]]
```

```{r}
k <- 1
xmltop[[k]][["drugbank-id"]] #The drug's ID is DB00001
xmltop[[k]][["name"]] # The drug's name is Lepirudin
xmltop[[k]][["indication"]][[1]]#For what the drug is used
xmltop[[k]][["description"]]# Description of the drug 
xmltop[[k]][["snp-effects"]]#Target protein name, gene name, uniprot id, SNP id...
xmltop[[k]][["snp-adverse-drug-reactions"]]#No SNP adverse drug reaction
xmltop[[k]][["categories"]][[2]]#Category of the drug: Anticoagulants
xmltop[[k]][["targets"]][[1]][[7]][["gene-name"]]#Gene of the target protein
xmltop[[k]][["carriers"]][[1]][[7]][["gene-name"]]
```

### 2) Extracting information:
*Dr Sina Rueger kindly provided me the code for a function that extracts data from the xml file :
```{r}
extract.xml <- function(string, data){
  tmp <- data[[string]]
  if(xmlSize(tmp)==0)
  {
    return(NA)
  }
  #Return NA if the string argument is empty
  if (xmlSize(tmp) == 1 & !(string %in% c("transporters","targets","enzymes","carriers")))
    #If the size of string equals 1, then return tmp[[1]] as a character 
  {
    return(as(tmp[[1]], "character"))
  }else{
    if (string == "categories")
    {
      tmp2 <- sapply(1:xmlSize(tmp), function(x) as(tmp[[x]][["category"]][[1]],
                                                    "character"))
      
      tmp.out <- paste0(tmp2, collapse = "/")
      return(tmp.out)
    }else{
      if (string %in% c("transporters","targets","enzymes","carriers"))
      {
        tmp2 <- sapply(1:xmlSize(tmp), function(x)
                   {
                       gene <- tmp[[x]][[7]][["gene-name"]][[1]]
                       if (!is.null(gene))
                       {
                           as(gene,"character")
                       }else{
                           NA
                       }
                   }
                       ) ## maybe more???
        ##paste
        tmp.out <- paste0(na.omit(tmp2), collapse = "/")
        return(tmp.out)
      
        ## if larger and not present, then NA
      }else{
          return(NA)
      }
    }
  }
  
}
```

#Now test the function yourself:

```{r}
k <- 5443
data <- xmltop[[k]]
extract.xml("drugbank-id", data)
extract.xml("name", data)
extract.xml("indication", data)
extract.xml("description", data)
extract.xml("snp-effects", data)
extract.xml("snp-adverse-drug-reactions", data)
extract.xml("categories", data)
extract.xml("targets", data)
extract.xml("enzymes", data)
extract.xml("carriers", data)
extract.xml("transporters", data)
```
--> As you can see extract.xml extracts the name of the information you seek as a string of characters without the tree structure of xml (more humanly readable).
# A function that stores the info in a data frame :
```{r}
xml2df <- function(strings, data)
{
   out <- data.frame(t(unlist(sapply(strings, function(x) extract.xml(x, data)))))
            #Apply the extract.xml function on the whole data 
            #the output will be stored in a data frame called "out"
                         
  return(out)
  
}
```

# Now test this function for a Drug : 
```{r}
strings <- c("drugbank-id", "name", "indication", "description",
             "snp-effects","snp-adverse-drug-reactions" ,"categories",
             "targets","transporters","enzymes","carriers")
k <- 5443 ## which(DB.raw$name == "Atenolol")
data <- xmltop[[k]]

xml2df(strings,data)
```

```{r}
library(parallel)
tmp <- mclapply(1:xmlSize(xmltop), function(k)
  {
     out <-  xml2df(strings,xmltop[[k]])
    return(out) 
  }, mc.cores = 1)
```

tmp is a large list of 11292 elements, convert this into a data frame :
```{r}
#install.packages('data.table')
library(data.table)
DB <- rbindlist(tmp)
#install.packages("dplyr")
library(dplyr)
DB <- (DB %>% mutate_if(is.factor,as.character) )#Turn all factors into characters
```
Save your precious filtered data:
```{r}
save(DB,file="DrugBank_data.RData")
```

